---
title: "Multiple linear regression"
author: "Wing-Ho Ko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rgl)
library(car)
library(lmtest)
library(palmerpenguins)
```

### Additive case

```{r}
penguins_sim <- penguins |> select(species:body_mass_g) |> drop_na()
```

```{r}
penguins_sim
```

Regress body mass on bill length and bill depth

```{r}
pg_lm_2 <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm, penguins_sim)
```

```{r}
summary(pg_lm_2)
```

Check normality and heteroscedasticity as before

```{r}
bptest(pg_lm_2)
```

```{r}
shapiro.test(residuals(pg_lm_2))
```

Visualize how the variance may be skewed

```{r}
pg_fitted_2 = data.frame(fit=fitted(pg_lm_2), resid=residuals(pg_lm_2))

ggplot(pg_fitted_2, aes(x=fit, y=resid)) + geom_point()
```

Check the span of values in the explanatory variables:

```{r}
summary(penguins_sim)
```

Create new data for prediction

Note the use of the `rep` function, once with the argument `times` and the other with the argument `each`

```{r}
new_blength = seq(32, 60, 1)
new_bdepth = seq(13, 22, 1)

penguins_new <- data.frame(
  bill_length_mm = rep(new_blength, times=length(new_bdepth)),
  bill_depth_mm = rep(new_bdepth, each=length(new_blength))
)
```

```{r}
penguins_new
```

Alternatively, use the `cart_expand()` function in the included source file `cart_expand.R`

```{r}
source("cart_expand.R")

cart_expand(
  seq(32, 60, 1), seq(13, 22, 1),
  names = c("bill_length_mm", "bill_depth_mm")
)
```

Now making the prediction

```{r}
penguins_pred_2 <- penguins_new |> mutate(
  pred = predict(pg_lm_2, penguins_new)
)

penguins_pred_2
```

Visualize the prediction using `rgl`

```{r}
pred_mat_2 <- as.matrix(penguins_pred_2$pred)
dim(pred_mat_2) <- c(length(new_blength), length(new_bdepth))

persp3d(new_blength, new_bdepth, pred_mat_2, front = "lines", back = "lines")
```

### Non-additive case

Again use the penguins dataset

```{r}
penguins_sim <- penguins |> select(species:body_mass_g) |> drop_na()
```

```{r}
penguins_sim
```

Non-additive model on the effect of bill length and bill depth on body mass:

```{r}
pg_lm_2x <- lm(body_mass_g ~ bill_length_mm * bill_depth_mm, penguins_sim)
```

```{r}
summary(pg_lm_2x)
```

Calculating confidence interval for the estimated parameters

```{r}
confint(pg_lm_2x)
```

Construct the grid to make prediction on

```{r}
source("cart_expand.R")

new_blength <- seq(32, 60, 1)
new_bdepth <- seq(13, 22, 1)

penguins_new <- cart_expand(
  new_blength, new_bdepth,
  names = c("bill_length_mm", "bill_depth_mm")
)
```

Making prediction:

```{r}
penguins_pred_2x <- penguins_new |> mutate(
  pred = predict(pg_lm_2x, penguins_new)
)

penguins_pred_2x
```

Visualize the prediction using `rgl`

```{r}
pred_mat_2x <- as.matrix(penguins_pred_2x$pred)
dim(pred_mat_2x) <- c(length(new_blength), length(new_bdepth))

persp3d(new_blength, new_bdepth, pred_mat_2x, front = "lines", back = "lines")
```

Compare the additive and interacting models using AIC and BIC

```{r}
AIC(pg_lm_2, pg_lm_2x)
```

```{r}
BIC(pg_lm_2, pg_lm_2x)
```

Try to fit body mass using bill length, bill depth, and flipper length as explanatory variables. Consider the additive, non-additive second order, and fully interacting model

```{r}
pg_lm_3 = lm(
  body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm, 
  penguins_sim
)
```

```{r}
pg_lm_3x = lm(
  body_mass_g ~ 
    bill_length_mm + bill_depth_mm + flipper_length_mm + 
    bill_length_mm:bill_depth_mm + 
    bill_length_mm:flipper_length_mm + 
    bill_depth_mm:flipper_length_mm, 
  penguins_sim
)
```

```{r}
pg_lm_3xx = lm(
  body_mass_g ~ bill_length_mm * bill_depth_mm * flipper_length_mm, 
  penguins_sim
)
```

Check the summary for each of them.

```{r}
summary(pg_lm_3)
```

```{r}
summary(pg_lm_3x)
```

```{r}
summary(pg_lm_3xx)
```

Check AIC comparison:

```{r}
AIC(pg_lm_3, pg_lm_3x, pg_lm_3xx)
```

Do the same for BIC:

```{r}
BIC(pg_lm_3, pg_lm_3x, pg_lm_3xx)
```

Start with additive model then try to add terms one-by-one:

```{r}
step(
  pg_lm_3, 
  body_mass_g ~ bill_length_mm * bill_depth_mm * flipper_length_mm,
  direction="forward"
)
```

Start with the fully interacting model, then try to subtract term one-by-one:

```{r}
step(pg_lm_3xx, direction="backward")
```

