---
title: "Simple Linear Regression"
author: "Wing-Ho Ko"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(janitor)
library(tidyverse)
library(lmtest)
library(palmerpenguins)
```

## Simple linear regression

### Loading data

Load and simplify the penguins dataset

```{r}
penguins_sim <- penguins |> select(species:body_mass_g) |> drop_na()
```

```{r}
penguins_sim
```

### Perform simple linear regression

Perform linear regression with body mass being the response variable and bill length being the explanatory variable:

```{r}
pg_lm <- lm(body_mass_g ~ bill_length_mm, penguins_sim)
```

### Extracting information from the fit

Printout useful summary of the model

```{r}
pg_sm <- summary(pg_lm)
print(pg_sm)
```

Extract the estimated standard deviation programmatically

```{r}
pg_sm$sigma
```

Extract the R^2 value programmatically

```{r}
pg_sm$r.squared
```

Extract the coefficient of the linear model programmatically

```{r}
coef(pg_lm)
```

Extract the residual of the linear fit

```{r}
residuals(pg_lm)
```

Plot the original data point against the fitted values

```{r}
penguins_fitted <- penguins_sim |> mutate(fit = fitted(pg_lm))

ggplot(penguins_fitted) + 
  geom_point(aes(x=bill_length_mm, y=body_mass_g)) +
  geom_line(aes(x=bill_length_mm, y=fit), color="blue", linetype="dashed")
```

Plot the residuals against the value of the explanatory variable

```{r}
resid_df <- data.frame(
  bill_length_mm = penguins_sim$bill_length_mm,
  residual = residuals(pg_lm)
)

ggplot(resid_df, aes(x=bill_length_mm, y=residual)) + geom_point()
```

### Testing assumptions

Check for heteroskedasticity (same variance for all values of x) more quantitatively using `bptest()` from the `lmtest` package

```{r}
bptest(pg_lm)
```

While we're at it, test if the residual is normally distributed

```{r}
shapiro.test(residuals(pg_lm))
```

### Making new predictions

First, create a dataframe for new data:

```{r}
penguins_new <- data.frame(
  bill_length_mm = seq(30, 60, 0.2)
)
```

Making prediction using the fitted model

```{r}
predict(pg_lm, penguins_new)
```

Include the confidence interval

```{r}
pg_conf <- predict(pg_lm, penguins_new, interval="confidence")
pg_conf
```

Combine with the explanatory variable to form a dataframe:

```{r}
penguins_conf <- cbind(penguins_new, as.data.frame(pg_conf))
penguins_conf
```

Plot confidence interval with data:

```{r}
ggplot(penguins_conf, aes(x=bill_length_mm, y=fit)) + 
  geom_line(color="blue", linetype="dashed") + 
  geom_ribbon(aes(ymin=lwr, ymax=upr), fill="cyan", alpha=0.2) + 
  geom_point(aes(x=bill_length_mm, y=body_mass_g), penguins_sim) + 
  coord_cartesian(xlim=c(30, 60))
```

Compare with `geom_smooth()` with `method=lm`:

```{r}
ggplot(penguins_sim, aes(x=bill_length_mm, y=body_mass_g)) + 
  geom_smooth(method="lm") + geom_point() + 
  coord_cartesian(xlim=c(30, 60))
```

Do the same for prediction interval

```{r}
pg_pred <- predict(pg_lm, penguins_new, interval="prediction")
penguins_pred <- cbind(penguins_new, as.data.frame(pg_pred))
print(penguins_pred)
```

Now plot the prediction interval instead:

```{r}
ggplot(penguins_pred, aes(x=bill_length_mm, y=fit)) + 
  geom_line(color="forestgreen", linetype="dashed") + 
  geom_ribbon(aes(ymin=lwr, ymax=upr), fill="lawngreen", alpha=0.2) + 
  geom_point(aes(x=bill_length_mm, y=body_mass_g), penguins_sim) + 
  coord_cartesian(xlim=c(30, 60))
```
